---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
load("~/Desktop/GitHub/Peer-Rev-Assignment/brfss2013.Rdata")
```



* * *

## Part 1: Data

#Subset data into mental health data and sleep data for analysis. 
```{r}
brfss_mh_sl<-subset(brfss2013,select = c(X_state,employ1,hlthpln1,genhlth,menthlth,sleptim1))
```


* * *

## Part 2: Research questions

**Research quesion 1:**
#Which states generally show the highest number of days of poor mental health and the lowest average sleep times? 
#___________________________________________________________________________#

#1: Convert mental health scores from integers to factors for atomic manipulation.
```{r}
as.factor(brfss_mh_sl$menthlth)
```

#2: Create a new table with average mental health scores (number of days reported with poor mental health) and standard deviations: 

```{r}
state_avgs<-brfss_mh_sl%>%
  group_by(X_state)%>%
  summarise(mean_mh=mean(menthlth,na.rm = TRUE),sd_mh=sd(menthlth,na.rm=TRUE),mean_sleep=mean(sleptim1,na.rm=TRUE),sd_sleep=sd(sleptim1,na.rm=TRUE))
```

  
#Remove all non-state values 
```{r}
state_avgs<-state_avgs%>%
   filter(X_state!="Guam",X_state!="Puerto Rico",X_state!=0,X_state!=80)
 View(state_avgs)
 
 #Load tidyverse and urbnmapr for geocoding 
 
 library(tidyverse)
 library(urbnmapr)


```



#3: Use "states" and "statedata" to merge brfss data with map shapefile and coordinates 
```{r}
View(states)
View(statedata)
```


#Mutate state average data to include state FIPS codes from "states"
```{r}
state_avgs<-state_avgs%>%
  mutate(state_fips=statedata$state_fips)
```
  
#Create a map shapefile for visualization 
```{r}
mh_states_map<-left_join(state_avgs,states,by="state_fips")

##MENTAL HEALTH
mh_states_map %>%
  ggplot(aes(long, lat, group = group, fill = mean_mh)) +
  geom_polygon(color = NA) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "mean_mh")+ggtitle("Days of Poor Mental Health Reported by State")

```
```{r}
##SLEEP TIMES
mh_states_map %>%
  ggplot(aes(long, lat, group = group, fill = mean_sleep)) +
  geom_polygon(color = NA) +
  coord_map(projection = "albers", lat0 = 39, lat1 = 45) +
  labs(fill = "mean_sleep")+ggtitle("Average Reported Hours of Sleep by State")
```

##ANALYSIS: Based on the visualization, states showing higher numbers of days experiencing poor mental health are also states whose average sleep times are generally lower. 

**Research quesion 2:**
#Is there a statistical relationship between average hours of sleep and average days of poor mental health as indicated by the graph? 
#_________________________________________________________________________________________#

#Create a plot to visualize relationship, Sleep as the assumed independent variable. 

```{r}
scatter.smooth(x=state_avgs$mean_sleep,y=state_avgs$mean_mh,xlab = "Average Hours of Sleep Per State",ylab = "Number of Days of Low Mental Health")
```

#There appears to be a relationship showing fewer poor mental health days correlating with higher average hours of sleep. 

#STATISTICAL TEST: 
```{r}
cor(state_avgs$mean_sleep,state_avgs$mean_mh)
```

##This indicates a moderate negative relationsip between sleep and mental health days, suggesting that state populations who sleep more have better mental health. 

**Research quesion 3:**

```{r}
#Identify three lowest scoring states: 

#Top three pulled from table: Alabama, West Virginia, Kentucky. 

depression<-subset(brfss2013,select = c(X_state,addepev2))

table(depression$addepev2)
```
```{r}
95778/393707

```

#24% of total respondents had been told they had depression. 

#Pull only Alabama, West Virginia, and Kentucky 

```{r}
depression_alkewv<-subset(filter(depression,X_state=="West Virginia"|X_state=="Kentucky"|X_state=="Alabama"))

table(depression_alkewv$addepev2)


```
```{r}
5248/17932
```

```{r}
#29% of respondents in lower scoring states reported being told that they have depression. This is 4% higher than the percentage reported overall. 
```



